Type-Safe Modular Parsing

### 2017-09-04 SLE_17/sections/CaseStudy.tex
shortening.
Shorten more.

\subsection{Implementation}\label{subsec:implementation}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{resources/depGraph.pdf}
    \caption{Dependency graph of all calculi and components. Grey boxes are calculi; white boxes are components.}
    \label{fig:dependency}
\end{figure*}

@@

\subsection{Comparison}\label{subsec:cs-comparison}

@@

\paragraph{Standard of Comparison}
In the SLOC comparison, all blank lines and comments are excluded,
and we formatted the code of both implementations to guarantee that
the length of each line does not exceed 120 characters. Furthermore,
because \ilyaimpl{} has extra code such as semantics,
we removed all irrelevant code and only keep abstract
syntax definition, parser and pretty-printer for each calculus, to
ensure a fair comparison.

@@

\begin{table*}
    \centering
    \resizebox{0.7\textheight}{!}{
    \begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|}
      \hline
        \multirow{2}{*}{\bfseries Calculus Name} & \multicolumn{3}{ c| }{\bfseries SLOC} & \multicolumn{7}{ c| }{\bfseries Time (ms)} \\ \cline{2-11}
        \multicolumn{1}{|c|}{} & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyalongest{} & \bfseries (+/-)\% & \ourclass{} & \bfseries (+/-)\% \\
      \hline
      \input{resources/CaseStudyTable.tex}
      \hline
      \multicolumn{7}{c}{}
    \end{tabular}}
    \caption{Comparison of SLOC and execution time.}
    \label{tab:comparison}
    \vspace{-0.3cm}
\end{table*}

\paragraph{Comparison Results}
Table \ref{tab:comparison} shows results of the comparison.
Let us only check \ourimpl{} and \ilyaimpl{} for now.
The overall result is that 69.2\% of code is reduced using our
approach, and our implementation is 42.7\% slower.

@@

The right part of Table \ref{tab:comparison} suggests that the difference of running time between
using Object Algebras and class is little, roughly 1\%.
The usage of longest match combinator slows the performance by 7\%. The main reason of slower
execution time may be the overall structure of the modular parsing approach, because we indeed have
more intermediate function calls and method overriding. However, it is worth mentioning that
because of the memoization technique of Packrat parsers, we are only constant times
slower, the algorithmic complexity is still the same.
Since the slowdown seems to be caused by extra method
dispatching, in future work we wish to investigate techniques such as 
partial evaluation or meta-programming to eliminate such cost. The
work by B{\'e}guet and Manohar~\cite{Beguet:2014} is an interesting starting point.

>>>>

\vspace{-3pt}
\subsection{Implementation}\label{subsec:implementation}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.7\textwidth]{resources/depGraph.pdf}
    \caption{Dependency graph of all calculi and components. Grey boxes are calculi; white boxes are components.}
    \label{fig:dependency}
\end{figure*}

@@

\vspace{-3pt}
\subsection{Comparison}\label{subsec:cs-comparison}

@@

\paragraph{Standard of Comparison}
In terms of SLOC, all blank lines and comments are excluded,
and we formatted the code of both implementations to ensure that
the length of each line does not exceed 120 characters. Furthermore,
because \ilyaimpl{} has extra code like semantics,
we removed all irrelevant code, only kept abstract
syntax definition, parser and pretty-printer for each calculus, to
ensure a fair comparison.

@@

\begin{table*}\tiny
    \centering
    \resizebox{0.7\textheight}{!}{
    \begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|}
      \hline
        \multirow{2}{*}{\bfseries Calculus Name} & \multicolumn{3}{ c| }{\bfseries SLOC} & \multicolumn{7}{ c| }{\bfseries Time (ms)} \\ \cline{2-11}
        \multicolumn{1}{|c|}{} & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyaimpl{} & \ourimpl{} & \bfseries (+/-)\% & \ilyalongest{} & \bfseries (+/-)\% & \ourclass{} & \bfseries (+/-)\% \\
      \hline
      \input{resources/CaseStudyTable.tex}
      \hline
      \multicolumn{7}{c}{}
    \end{tabular}}
 	\vspace{-5pt}
    \caption{Comparison of SLOC and execution time.}
    \label{tab:comparison}
    \vspace{-0.3cm}
\end{table*}

\vspace{-4pt}
\paragraph{Comparison Results}
Table \ref{tab:comparison} shows results of the comparison.
Let us only check \ourimpl{} and \ilyaimpl{} for now.
The overall result is that 69.2\% of code is reduced using our
approach, and our implementation is 42.7\% slower.

@@

The right part of Table \ref{tab:comparison} suggests that the difference of running time between
using Object Algebras and class is little, roughly 1\%.
The use of longest match combinator slows the performance by 7\%. The main reason of slower
execution time may be the overall structure of the modular parsing approach, because we indeed have
more intermediate function calls and method overriding. However, it is worth mentioning that
because of the memoization technique of Packrat parsers, we are only constant times
slower, the algorithmic complexity is still the same.
Since the slowdown seems to be caused by extra method
dispatching, in future work we wish to investigate techniques like
partial evaluation or meta-programming to eliminate such cost. The
work by B{\'e}guet and Manohar~\cite{Beguet:2014} is an interesting starting point.

### 2017-09-04 SLE_17/sections/RelatedWork.tex
Shorten more.

\paragraph{Syntactically Extensible Parsing}
Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Viera2012,Warth2016,schwerdfeger09}
are a mainstream area of modular syntax and parsing. They allow users to write
modular grammars, where new
non-terminals and production rules can be introduced, some can even
override existing rules in the old grammar modules. For instance,
\textit{Rats!}~\cite{Grimm2006} constructs its own module system for
the collection of grammars, while NOA~\cite{Gouseti2014} uses Java
annotation to collect all information before producing an ANTLR~\cite{antlr1995} grammar
and the parsing code. Those parser
generators focus on the \textit{syntactic extensibility} of grammars:
they rely on whole compilation to generate a global parser, even if
there is only a slight modification in the grammar. Some of those
parser generators may statically check the correctness and unambiguity
of grammars. In contrast, because our approach is based on parser
combinators, there is no support for ambiguity checking.  However, as
far as we are aware, no extensible parser generators support separate
compilation or modular type-checking. It is worth mentioning that in~\cite{Viera2012}, users can define
grammar fragments as typed Haskell values, and combine them on the fly. Later they are processed
by a typed parser generator. Nevertheless this requires
a lot of advanced language features, making client complex. Our
approach is simple and a straightforward use of OO programming, 
and makes parsing code directly reusable.

@@

Besides, type-specific languages~\cite{omar14} do support
separate compilation and type safety. The modularity of parsing is incorporated in its type
system, which allows users to introduce new syntax for parsing. Whereas in our approach, parsing 
strategy is based on an external parsing library.
Our approach does not rely on an elaborated semantics extension, but simply uses a pattern
that is applicable in mainstream OO languages, where the type system ensures
type safety.

@@

\vspace{-3pt}
\paragraph{Parser Combinators} Parser combinators have become more and more
popular since~\cite{burge1975,Wadler1985}. Many parsing libraries produce recursive descent
parsers by introducing functional monadic
parser combinators~\cite{nott237}. Parsec~\cite{Leijen2001} is
perhaps the most popular parser combinator library in this line.
It is widely used in Haskell (with various ``clones'' in other languages)
for context-sensitive grammars with infinite lookahead. Nevertheless,
Parsec users suffer from manual left-recursion elimination,
high cost for backtracking and longest match composition issues,
as we discussed in Section~\ref{subsec:challenges}. Those limitations make Parsec
(and similar parsing techniques) inadequate for modular parsing.

@@

\vspace{-3pt}
\paragraph{Extensibility} Various design patterns~\cite{gamma1995design} in multiple
languages, have been proposed over the years to address extensibility
problems, such as the Expression Problem~\cite{wadler1998expression}.
The famous ``Datatypes \`a
la Carte'' (DTC)~\cite{swierstra2008data} approach represents modular ASTs using co-products
of every two functors. Several variants of DTC have been later proposed~\cite{Bahr2011,Bahr2014,Oliveira2015}.
All of that work essentially covers how to traverse and
consume extensible ASTs. However they do not
address the problem of \emph{modularly
parsing extensible ASTs}. Only in Bahr's~\cite{Bahr2011} work \emph{unfolds} is briefly mentioned,
yet it does not cover parsing.

@@

It is worth mentioning that Scala case classes~\cite{emir2007matching} provide a near
solution to the Expression Problem. Case classes can be
modularly added, but they do not enforce
exhaustiveness of pattern matching for extensible operations. In other
words run-time pattern matching errors can happen when writing extensible code with case classes. So that full static type-safety is not ensured. Nevertheless case classes are a pragmatic approach, which is
widely used in practice. Therefore, modular parsing techniques may be
of value for extensible code using case classes.  The approach we
presented in Section~\ref{sec:inheritance} can readily be adapted to case classes:
all that the users need to do is to use case classes, instead of
standard OO classes in their code.
\vspace{-5pt}

>>>>

\vspace{-4pt}
\paragraph{Syntactically Extensible Parsing}
Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Viera2012,Warth2016,schwerdfeger09}
are a mainstream area of modular syntax and parsing. They allow users to write
modular grammars, where new
non-terminals and production rules can be introduced, some can even
override existing rules in the old grammar modules. For instance,
\textit{Rats!}~\cite{Grimm2006} constructs its own module system for
the collection of grammars, while NOA~\cite{Gouseti2014} uses Java
annotation to collect all information before producing an ANTLR~\cite{antlr1995} grammar
and the parsing code. Those parser
generators focus on the \textit{syntactic extensibility} of grammars:
they rely on whole compilation to generate a global parser, even if
there is only a slight modification in the grammar. Some of those
parser generators may statically check the correctness and unambiguity
of grammars. In contrast, because our approach is based on parser
combinators, there is no support for ambiguity checking.  However, as
far as we are aware, no extensible parser generators support separate
compilation or modular type-checking. It is worth mentioning that in~\cite{Viera2012}, users can define
grammar fragments as typed Haskell values, and combine them on the fly. Later they are processed
by a typed parser generator. Nevertheless this requires
a lot of advanced language features, making client complex. Our
approach is simple and a straightforward use of OO programming, 
and makes parsing code directly reusable.

@@

Besides, type-specific languages~\cite{omar14} do support
separate compilation and type safety. The modularity of parsing is incorporated in its elaborated type
system. Whereas our approach does not rely on an extension to semantics, but simply uses a pattern
that is applicable in mainstream OO languages, where the type system ensures
type safety.

@@

\vspace{-5pt}
\paragraph{Parser Combinators} Parser combinators have become more and more
popular since~\cite{burge1975,Wadler1985}. Many parsing libraries produce recursive descent
parsers by introducing functional monadic
parser combinators~\cite{nott237}. Parsec~\cite{Leijen2001} is
perhaps the most popular parser combinator library in this line.
It is widely used in Haskell (with various ``clones'' in other languages)
for context-sensitive grammars with infinite lookahead. Nevertheless,
Parsec users suffer from manual left-recursion elimination,
high cost for backtracking and longest match composition issues,
as we discussed in Section~\ref{subsec:challenges}. Those limitations make Parsec
(and similar parsing techniques) inadequate for modular parsing.

@@

\vspace{-5pt}
\paragraph{Extensibility} Various design patterns~\cite{gamma1995design} in multiple
languages, have been proposed over the years to address extensibility
problems, such as the Expression Problem~\cite{wadler1998expression}.
The famous ``Datatypes \`a
la Carte'' (DTC)~\cite{swierstra2008data} approach represents modular ASTs using co-products
of every two functors. Several variants of DTC have been later proposed~\cite{Bahr2011,Bahr2014,Oliveira2015}.
All of that work essentially covers how to traverse and
consume extensible ASTs. However they do not
address the problem of \emph{modularly
parsing extensible ASTs}. Only in Bahr's~\cite{Bahr2011} work \emph{unfolds} is briefly mentioned,
yet it does not cover parsing.

@@

It is worth mentioning that Scala case classes~\cite{emir2007matching} provide a near
solution to the Expression Problem. Case classes can be
modularly added, but they do not enforce
exhaustiveness of pattern matching for extensible operations. In other
words run-time pattern matching errors can happen when writing extensible code with case classes. So that full static type-safety is not ensured. Nevertheless case classes are a pragmatic approach, which is
widely used in practice. Therefore, modular parsing techniques may be
of value for extensible code using case classes.  The approach we
presented in Section~\ref{sec:inheritance} can readily be adapted to case classes:
all that the users need to do is to use case classes, instead of
standard OO classes in their code.
\vspace{-7pt}
